Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2020-11-12 mice_synthesizing/paper.Rmd
add key phrase
Add content + outline to intro.

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general thought is to use actually observed data to generated multiply imputed synthetic datasets that can be freely disclosed. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed data set as a whole with multiple synthetic versions. Additionally, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available data sets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, say, income values above a certain threshold.

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation for missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Doing this multiple times would allow for drawing correct statistical inferences, similar to those obtained from ordinary multiple imputation for incomplete data sets. The results of the multiply synthetisized data sets should be pooled into an single inference, so that the researcher can draw valid inference from a single pooled analysis. To that respect, the variance should reflect the added variability that is induced by the imputation procedure.

@@

Since multiple imputation of synthetic data is closely linked to multiple imputation for missing data, using existing missing data software that research are familiar with already, would dramatically reduce the effort of creating synthetic data. The R-package `mice` [@mice] implements multiple imputation for missing data in a straightforward and user-friendly manner, but is not restricted to the imputation of missing data. Since `mice` can be easily adjusted to impute all observed values, or a subset hereof, we propose that it is able to ease the creation of synthetic datasets.

>>>>

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general idea is to use actually observed data to generated multiply imputed synthetic datasets that can be freely disclosed. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed data set as a whole with multiple synthetic versions. Alternatively, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available data sets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, say, income values above a certain threshold.

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation of missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Similarly to multiple imputation of missing data, the multiple synthetic datasets allow for correct statistical inferences. The results of the multiply synthesized data sets should be pooled into a single inference, so that the researcher can draw valid inference from a single pooled analysis. To that respect, the variance should reflect the added variability that is induced by the imputation procedure.

@@

To minimize additional effort of creating synthetic datasets on behalf of the researcher, software aimed at multiple imputation of missing data can be employed. Especially if researchers used this software at an earlier stage in the research process, or acquired familiarity with it during earlier projects, the additional burden of creating synthetic datasets is relatively small. The R-package `mice` [@mice] implements multiple imputation of missing data in a straightforward and user-friendly manner. However, the functionality of `mice` is not restricted to the imputation of missing data, but allows to impute any value in the data. Thus, `mice` can be utilized for the creation of multiply imputed synthetic datasets.

After creating the multiply imputed synthetic data sets, the goal is to obtain valid statistical inferences from them. In the missing data framework, this is done by performing statistical analyses on all imputed data sets, and pooling the results of the analyses according to Rubin's rules [-@rubin_multiple_1987]. In the synthetic data framework, the same procedure is followed. However, valid statistical results can only be obtained under the premise that the statistical properties of the analyses are correct. Analyzing synthetic data specically requires appropriate variance estimates. Notably, the variance estimator that is suitable for analyzing multiply imputed datasets for missing data is inappropriate when multiply imputed synthetic datasets are analyzed. As a result, multiple variance estimators have been proposed to analyze multiply imputed synthetic datasets correctly, that differ according to whether the data consists entirely of synthetic values, or only a subset of the data is synthetic.

The remainder of this paper is organized as follows. First, we will shortly describe the `mice` algorithm for the creation of synthetic data that reassures the privacy and confidentiality of the participants. Second, we will demonstrate a straightforward workflow for imputation of synthetic data with `mice`. The third step is to summarize the variance estimators that have been proposed in the past. Finally, we will assess which variance estimator provides valid inferences from the synthetic data.

# Generating synthetic data with `mice`

KEY: Het moet niet mogelijk zijn om de echte waarden te reverse engineeren. De stochastische component van multiple imputation zorgt er voor dat dit onmogelijk is. Wat wel kan worden afgeleid is het ware data genererende model - en dat is nou precies de bedoeling.
